Handling missing data in smartphone location logs

### 2017-11-30 thesisReport.Rmd
added pdf
added lables to accuracy plot, changed Y variable

These obstacles can be overcome by using objective data on human mobility. Such data  can now be obtained using the Global Positioning System (GPS). GPS uses the distance between a device and several satellites to determine location. Within behavioural sciences, researchers have used GPS data to investigate wideranging topics. For example,  @zenk_how_2009 investigate the effects of the food environment on eating patterns. @harari_using_2016 look at the movement correlates of personality.  @wang_smartgpa:_2015  does the same for academic performance.  @palmius_detecting_2017 use mobility patterns to predict bipolar depression.

@@

Noisy data can lead to inaccurate conclusions if it is not accounted for. Suppose we wish to calculate an individual's movement in a day. A simple approach would be to calculate the sum of the distance between each measurement. But if there is noise, the coordinates will vary even though the individual is not moving. If the measurements are frequent and noisy, we will calculate a lot of movement, even if the individual did not move at all! The problem is further complicated because missing data and noisy measurements are related. Methods used by researchers to reduce noise, such as throwing out inaccurate measurements [e.g. @palmius_detecting_2017],  can exacerbate the severity of the missing data problem.

@@

```{r, accuracyPlot2, echo=FALSE, fig.cap="Measures of user activity and measurement accuracy on February 15th 2017.The upper chart shows the rolling mean distance from the next point in meters over the course of the day. The first peak corresponds to the first journey from the users home a gym around 8am. The second, smaller peak before 10 reflects a journey from the gym to the lecture theatre. Both journies can be seen in Figure 1. The lower chart shows the rolling mean accuracy over the course of the day. In both cases, the rolling window is 3. The figure shows that measurement inaccuracy is sometimes related to the movement of the individual.",out.width = '100%'}
include_graphics("img/accuracyLocShift.png")

>>>>

These obstacles can be overcome by using objective data on human mobility. Such data  can now be obtained using the Global Positioning System (GPS). GPS uses the distance between a device and several satellites to determine location. GPS measurements can be used to infer a vast range of socioeconomic and behavioural measures, including where the individual lives, how much time he or she spends at home and where (and how) they travel.

Within behavioural sciences, researchers have used GPS data to investigate wideranging topics. For example,  @zenk_how_2009 investigate the effects of the food environment on eating patterns. @harari_using_2016 look at the movement correlates of personality, finding that extrovers individuals spend less time at home than introverts.  @wang_smartgpa:_2015  looks at how academic performance is affected by movement patterns.  @palmius_detecting_2017 use mobility patterns to predict bipolar depression.

@@

Noisy data can lead to inaccurate conclusions if it is not accounted for. Suppose we wish to calculate an individual's movement in a day. A simple approach would be to calculate the sum of the distance between each measurement. But if there is noise, the coordinates will vary even though the individual is not moving. If the measurements are frequent and noisy, we will calculate a lot of movement, even if the individual did not move at all! This issue is also visualised in Figure 1. The problem is further complicated because missing data and noisy measurements are related. Methods used by researchers to reduce noise, such as throwing out inaccurate measurements [e.g. @palmius_detecting_2017],  can exacerbate the severity of the missing data problem.

@@

```{r, accuracyPlot2, echo=FALSE, fig.cap="Measures of user activity and measurement accuracy on February 15th 2017.The upper chart shows the distance from the next measured point in meters over the course of the day. All journeys are marked with a red line. The first peak corresponds to the first journey from the user's home to a gym around 8am. The second, smaller peak before 10 reflects a journey from the gym to the nearby lecture theatre. Both journeys can be seen in Figure 1. All other journeys are not shown in Figure 1. The large jump between journey 5 and 6 is measurement error. The lower chart shows the accuracy over the course of the day. The figure shows that measurement inaccuracy is sometimes related to the movement of the individual.",out.width = '100%'}
include_graphics("img/accuracyLocShift.png")
